### 2015-05-25 docs/report/array_merging.tex

Once all the workers have executed the function provided to them on each element of their portion of the original array, they are to notify the main thread and provide the resulting \tabuffer{} or \tsabuffer{}.

@@

For the \tsabuffer{} case there are two different options:
\begin{enumerate}
  \item If all steps in the chain perform the \code{map} transformation, then no changes are required.
  \item If there is at least one \code{filter} step, the elements need to be placed at the beginning of the resulting \tstarray{} in order to avoid \textit{holes}.
\end{enumerate}

@@

The benchmark presented modifies \ttarray{} merge benchmarks to support \tstarray{}s. The implementation removes \dataview{} test case as it is not supported by \tstarray{}:
\bmcode{http://jsperf.com/sharedtypedarray-merge}{../../spikes/2.4.1/sharedMerge.js}

@@

\subsection{Conclusion}
The benchmark shows that the best alternative to merge small arrays into a larger one is using \code{Set} method of eighter \ttarray or \tstarray.

>>>>

Once all the workers have executed the \tfunction{} provided to them on each element of their portion of the original array, they are to notify the UI thread and provide the resulting \tabuffer{} or \tsabuffer{}.

@@

For the \tsabuffer{} case there are two possible situations:
\begin{enumerate}
  \item If all steps in the chain perform the \code{map} transformation, then no changes are required.
  \item If there is at least one \code{filter} step, the elements need to be placed at the beginning of the resulting \tstarray{} in order to avoid \textit{holes}.
\end{enumerate}

@@

The benchmark modifies the \ttarray{} merge benchmark to support instances of the \tstarray{} type. The implementation does not include the \dataview{} alternative as it is not supported by \tstarray{}s:
\bmcode{http://jsperf.com/sharedtypedarray-merge}{../../spikes/2.4.1/sharedMerge.js}

@@

\subsection{Conclusion}
The benchmark shows that the best alternative to merge small arrays into a larger one is using \code{Set} method of either \ttarray{} or \tstarray{}.

### 2015-05-25 docs/report/array_partitioning.tex

As explained in section~\ref{sec:serialization-and-transference} of this document, transferring \ttarray{} objects to and from workers requires the use of \code{Transferable}s to achieve acceptable performance. Since our library will receive \ttarray{} objects to be partitioned and handed over to multiple workers, it is important that it takes as little time as possible for the partitioning of the arrays.\footnote{This does not apply to \tstarray{}s as those can be transferred without being partitioned.}

If the library is configured to use \(N\) workers then the original data in the \ttarray{} will copied into \(N\) smaller \ttarray{}s. For ideal performance it would be nice to be able to pass a pointer into the array and an amount of elements to process, but \ttarray{}s do not support shared memory and element ownership is tranferred to the Web Workers, the elements must be copied to the sub-arrays.

\subsection{Copying \ttarray{}s}
\label{subsec:copying-typed-arrays}
Sections 22.2 and 24.1 of the ECMAScript 6 draft \cite{es6} specify the API for \ttarray{} and \tabuffer{} objects respectively. Some of the exposed operations provide a way to copy the contents of an \tabuffer{} into another \tabuffer{}, so they are to be considered as viable alternatives for the copy operation. Other alternatives to consider are:
\begin{itemize}
  \item Implementing a non-native function that copies the array elements one by one
  \item Using a \code{Blob} to store the contents of a \ttarray{} and reading the \code{Blob}'s contents into a new \ttarray{}.
\end{itemize}

@@

Based on those results we can discard the \code{Blob} approach for splitting \ttarray{}s.

@@

\subsubsection{\ttarray{} split benchmark}
Based on the result from the previous benchmark we created a new benchmark to understand which alternative is better to partition an array into parts:
\bmcode{http://jsperf.com/arraybuffer-split/3}{../../spikes/2.3.1/splitComparison.js}

The results for it are displayed in Figure~\ref{fig:typed-array-split}.
\img{Comparing \ttarray{} split approaches}{../../spikes/2.3.1/splitComparison}{fig:typed-array-split}

\subsubsection{Conclusion}
The benchmark shows that there are three very similar alternatives to partition a \ttarray{}. We are going to use \textit{Buffer subarray} as it yields better performance in a newer version of Chrome. The implementation of the slice function could look something like this:
\begin{lstlisting}[caption=Possible approach to copy a slice of a \ttarray{}]
function typedArraySlice(array, from, to) {
  return new array.constructor(array.subarray(from, to));
}
\end{lstlisting}

>>>>

As explained in Section~\ref{sec:serialization-and-transference} of this document, transferring \ttarray{} objects to and from Web Workers requires the use of \code{Transferable}s to achieve acceptable performance. Since our library will receive \ttarray{} objects to be partitioned and handed over to multiple workers, it is important that it spends as little time as possible partitioning the arrays.\footnote{This does not apply to \tstarray{}s as those can be transferred without being partitioned.}

If the library is configured to use \(N\) workers then the original data in the \ttarray{} will copied into \(N\) smaller \ttarray{}s. For ideal performance it would be nice to be able to pass a \textit{pointer} into the array and an amount of elements to process, but because \ttarray{}s do not support shared memory, and element ownership is tranferred to the Web Workers, the elements must be copied to the sub-arrays.

\subsection{Copying \ttarray{}s}
\label{subsec:copying-typed-arrays}
Sections 22.2 and 24.1 of the ECMAScript 6 draft \cite{es6} specify the API for \ttarray{} and \tabuffer{} objects respectively. Some of the exposed operations provide a way to copy the contents of an \tabuffer{} into another \tabuffer{}, so they are to be considered as viable alternatives for the copy operation. Other alternatives to consider are:
\begin{itemize}
  \item Implementing a non-native function that copies the array elements one by one.
  \item Using a \code{Blob} to store the contents of a \ttarray{} and reading the \code{Blob}'s contents into a new \ttarray{}.
\end{itemize}

@@

Based on those results we can discard the \code{Blob} approach for partitioning \ttarray{}s.

@@

\subsubsection{\ttarray{} partition benchmark}
Based on the result from the previous benchmark we created a new benchmark to understand which alternative is better to partition an array:
\bmcode{http://jsperf.com/arraybuffer-split/3}{../../spikes/2.3.1/splitComparison.js}

The results for it are displayed in Figure~\ref{fig:typed-array-split}.
\img{Comparing \ttarray{} partition approaches}{../../spikes/2.3.1/splitComparison}{fig:typed-array-split}

\subsubsection{Conclusion}
The benchmark shows that there are three very similar alternatives to partition a \ttarray{}. We decided to use \textit{Buffer subarray}. The implementation of the slice function could look something like this:
\begin{lstlisting}[caption=Possible approach to copy a slice of a \ttarray{}]
function typedArraySlice(array, from, to) {
  return new array.constructor(array.subarray(from, to));
}
\end{lstlisting}

### 2015-05-25 docs/report/code_sharing.tex

\subsection{Transferring strings}
Function's code not only needs to be serialized but also transferred. For that reason it was also worth comparing the time it takes to encode/decode a \tstring{} and tranfer the resulting \tabuffer{} to the worker against just passing the \tstring{} to the worker.

@@

\subsubsection{\tstring{} serialization and tranfer benchmarks}
We created a benchmark using two functions whose string representation has different lengths to see if the function's length affected the serialization and transfer time. We measured three approaches to transfer the string:
\begin{enumerate}
  \item Encoding the \tstring{} into an \tabuffer{}, sending it as a \code{Tranferable} and decoding it on the Web Worker.
  \item Sending the string directly
  \item Using a \code{Blob} to store the stream and retrieving it in the Worker.
\end{enumerate}

@@

\subsubsection{\tstring{} serialization and shared tranfer benchmark}

@@

\subsubsection{\tstring{} and \ttarray{} serialization and tranfer benchmarks}
We created a benchmark that tranfers a \ttarray{} and a \tstring{} that are properties of the same object back and forth between the main thread an a worker to understand if transferring additional objects affected the transfer time when comparing it with just tranferring a \tstring{}.

@@

\subsubsection{Conclusion}
As the benchmarks show, the fastest way to tranfer the \tstring{} between the UI thread and workers is by sending it directly.

>>>>

\subsection{Transferring strings}
Function's code not only needs to be serialized but also transferred. For that reason it was also worth comparing the time it takes to encode/decode a \tstring{} and transfer the resulting \tabuffer{} to the worker against just passing the \tstring{} to the worker.

@@

\subsubsection{\tstring{} serialization and transfer benchmarks}
We created a benchmark using two functions whose string representation has different lengths to see if the function's length affected the serialization and transfer time. We measured three approaches to transfer the string:
\begin{enumerate}
  \item Encoding the \tstring{} into an \tabuffer{}, sending it as a \code{Transferable} and decoding it on the Web Worker.
  \item Sending the string directly
  \item Using a \code{Blob} to store the stream and retrieving it in the Worker.
\end{enumerate}

@@

\subsubsection{\tstring{} serialization and shared transfer benchmark}

@@

\subsubsection{\tstring{} and \ttarray{} serialization and transfer benchmarks}
We created a benchmark that transfers a \ttarray{} and a \tstring{} that are properties of the same object back and forth between the main thread an a worker to understand if transferring additional objects affected the transfer time when comparing it with just transferring a \tstring{}.

@@

\subsubsection{Conclusion}
As the benchmarks show, the fastest way to transfer the \tstring{} between the UI thread and workers is by sending it directly.

### 2015-05-25 docs/report/inlining.tex
Benchmark link fixes

Which was implemented as shown in the following listing for the parallel case:
\begin{lstlisting}[caption=Sepia tone initial parallel implementation]
pjs(new Uint32Array(canvasData.data.buffer)).map(function(pixel){
  function noise() {
      return Math.random() * 0.5 + 0.5;
  };

@@

\subsection{The problem}
Taking the experience from previous experiments into account we considered that the performance of the two implementations with a source \code{Uint32Array} of about \(8E7\) (eight million) elements should be similar, even favoring the parallel one. The initial measurements showed that the linear implementation fared approximately four times better than the parallel one.

\subsection{The hypothesis}
As the two code samples were similar our hypothesis was that for some reason some v8 optimizations were not taking place. It was a definite possibility considering that the parallel implementation depends on the \tfunction{} constructor to create \tfunction{} objects in the Web Workers. Specifically we believed that the internal functions \code{noise}, \code{clamp} and \code{colorDistance} were not being inlined. Inlining could be the cause of both benefits\cite{inline-pro} and drawbacks\cite{inline-cons} but the belief was that performance was being affected because of lack of it

@@

\subsubsection{Parallel implementation generated code}
Figures \ref{fig:parallel-clamp} through \ref{fig:parallel-noise} are related to the serial implementation. As it can be seen from them, only \code{Math.random} is being inlined.
\img{\code{clamp} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-clamp}{fig:parallel-clamp}
\img{\code{colorDistance} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-colorDistance}{fig:parallel-colorDistance}
\img{\code{mapper} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-mapper}{fig:parallel-mapper}
\img{\code{noise} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-noise}{fig:parallel-noise}

@@

We created a benchmark to understand the difference between the initial version (the one in which functions were not inlined) and the final one (with inlining in charge of the developer):
\bmcode{http://jsperf.com/pjs-map-inlining}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining.js}

The results for it are displayed in Figure~\ref{fig:pjs-map-inlining}.
\img{Comparing string encoding/decoding approaches}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining}{fig:pjs-map-inlining}

>>>>

It was implemented as shown in the following listing for the parallel case:
\begin{lstlisting}[caption=Sepia tone initial parallel implementation]
pjs(new Uint32Array(canvasData.data.buffer)).map(function(pixel){
  function noise() {
      return Math.random() * 0.5 + 0.5;
  };

@@

\subsection{The problem}
Taking the experience from previous experiments into account we considered that the performance of the two implementations with a source \code{Uint32Array} of about 8,000,000 (eight million) elements should be similar, even favoring the parallel one. The initial measurements showed that the linear implementation fared approximately four times better than the parallel one.

\subsection{The hypothesis}
As the two code samples were similar our hypothesis was that for some reason some v8 optimizations were not taking place. It was a definite possibility considering that the parallel implementation depends on the \tfunction{} constructor to create \tfunction{} objects in the Web Workers. Specifically we believed that the internal functions \code{noise}, \code{clamp} and \code{colorDistance} were not being inlined. Inlining could be the cause of both benefits\cite{inline-pro} and drawbacks\cite{inline-cons} but the belief was that performance was being affected because of lack of it.

@@

\subsubsection{Parallel implementation generated code}
Figures \ref{fig:parallel-clamp} through \ref{fig:parallel-noise} are related to the parallel implementation. As it can be seen from them, only \code{Math.random} is being inlined.
\img{\code{clamp} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-clamp}{fig:parallel-clamp}
\img{\code{colorDistance} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-colorDistance}{fig:parallel-colorDistance}
\img{\code{mapper} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-mapper}{fig:parallel-mapper}
\img{\code{noise} function - parallel implementation}{../../spikes/3.1.3/irhydra-pjs-noinlined/ww-noise}{fig:parallel-noise}

@@

We created a benchmark to understand the difference between the initial version (the one in which functions were not inlined) and the final one (with inlining in charge of the developer):
\bmcode{http://jsperf.com/pjs-map-inlining/6}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining.js}

The results for it are displayed in Figure~\ref{fig:pjs-map-inlining}.
\img{Comparing a version of \code{map} without inlining against a developer inlined one}{../../spikes/3.1.3/pjsMapInlining/pjsMapInlining}{fig:pjs-map-inlining}

### 2015-05-25 docs/report/modules.tex

The two most known module formats/API in JavaScript are known as:
\begin{description}
\item[CommonJS] Recognized for being used in node.js~\cite{nodejs}.
\item[AMD~\cite{amd}] The most known implementation being require.js~\cite{requirejs}.
\end{description}

@@

Instead of making the choice for users, our library will support both systems in a way that makes it easier for users to consume either of them.

@@

\subsection{AMD support}
This example shows how to use the library with require.js. It assumes a folder structure like the one shown below:
\begin{lstlisting}[language=bash, caption=AMD folder structure]
.
|__ index.html
|__ scripts
    |__ main.js
    |__ require.js
\end{lstlisting}

@@

\subsection{Conclusion}
As the previous examples show, using our library with either module system is really simple and is done in a way that should feel familiar for users of those module formats.

>>>>

The two most known module formats/API in JavaScript are:
\begin{description}
\item[CommonJS] Recognized for being used in node.js~\cite{nodejs}.
\item[AMD~\cite{amd}] The most known implementation being require.js~\cite{requirejs}.
\end{description}

@@

Instead of making the choice for users, our library will support both systems in a way that makes it easy for users to consume either of them.

@@

\subsection{AMD support}
This example shows how to use the library with require.js. It assumes a folder structure like the one in the following listing:
\begin{lstlisting}[language=bash, caption=AMD folder structure]
.
|__ index.html
|__ scripts
    |__ main.js
    |__ require.js
\end{lstlisting}

@@

\subsection{Conclusion}
As the previous examples show, using our library with either module system is really simple and is done in a way that should feel familiar for users of those systems.

### 2015-05-25 docs/report/serialization_transference.tex

\subsubsection{TypedArray With 100,000 elements}
The following is the code for the benchmark:
\bmcode{http://jsperf.com/transferrable-vs-cloning}{../../spikes/1.2/transferrableVsCloning.js}

@@

\subsubsection{TypedArray With 1,000,000 elements}
The following is the code for the benchmark:
\bmcode{http://jsperf.com/longer-transferrable-vs-cloning}{../../spikes/1.2/longerTransferrableVsCloning.js}

>>>>

\subsubsection{TypedArray With 100,000 elements}
The following is the code for the benchmark:
\bmcode{http://jsperf.com/transferrable-vs-cloning/3}{../../spikes/1.2/transferrableVsCloning.js}

@@

\subsubsection{TypedArray With 1,000,000 elements}
The following is the code for the benchmark:
\bmcode{http://jsperf.com/longer-transferrable-vs-cloning/3}{../../spikes/1.2/longerTransferrableVsCloning.js}
